---
title: "What is up !"
---
My name is Xingli Wang. I completed my Bachelor’s degree in Finance at Baruch College, where I built a solid foundation in financial theory, quantitative analysis, and decision-making. Currently, I am pursuing my Master’s degree in Business Analytics at Baruch College, with a concentration in data analytics. My academic interests include data-driven problem solving, statistical modeling, and applying analytical tools to support business strategies.

I look forward to contributing to our study group, exchanging knowledge, and learning from everyone’s perspectives.

```{r}
#| include: false
1+1
```
--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`

```{r}
#| echo: false
#| message: false
#| warning: false

if(!require("leaflet")){
    options(repos=c(CRAN="https://cloud.r-project.org"))
    install.packages("leaflet")
    stopifnot(require("leaflet"))
}

baruch_longitude <- -73.98333
baruch_latitude  <- +40.74028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom=17) |>
  addPopups(baruch_longitude, baruch_latitude, 
            "I am a Master's student at <b>Baruch College</b>!")
```