---
title: "MP01"
format: html
---



#### Upcoming Season of *Stranger Things*
"Stranger Things" Prepares for Final Bow After Dominating a Decade of Global Pop Culture

LOS GATOS, Calif. – As Netflix prepares to launch the fifth and final season of its flagship series, Stranger Things, at the end of 2025, new data confirms the show's unparalleled impact as a defining television phenomenon of the streaming era. Since its debut, the series has amassed a staggering 1.35 billion hours of viewership globally across its four seasons, making it one of the most-watched English-language series of all time. Its appeal is not only massive but enduring, with the series spending a remarkable 102 weeks in the Netflix Global Top 10, demonstrating a sustained popularity few shows can match. Furthermore, Stranger Things has proven to be a true multinational hit, consistently landing in the Top 10 in a record 93 different countries, from the United States to Brazil to South Korea. This global reach and longevity place Stranger Things in a league of its own, with a cultural footprint that rivals and often surpasses other modern classics, solidifying its legacy as more than just a show, but a global event



#### Commercial Success in India
Netflix Finds Its Hindi-Language Hit Formula, Unlocking Massive Growth in the Indian Market

MUMBAI, India – Netflix is celebrating a watershed moment in India, with its investment in original Hindi-language content driving unprecedented engagement and signaling immense growth potential in the world's most populous country. Blockbuster films like the action thriller Mission Majnu and the romantic drama Chor Nikal Ke Bhaga have become global sensations on the platform, despite being virtually unseen in the U.S. Top 10. This pattern highlights a content strategy tailored specifically for the Indian audience. By analyzing viewership, we can estimate the scale of this success; for instance, if all 69.9 million hours of Mission Majnu were viewed by Indian subscribers, it would suggest a massive and highly-engaged domestic customer base. The consistent charting of multiple Hindi-language originals each week points to a powerful long-term growth trend, demonstrating that local storytelling is the key to unlocking the vast Indian streaming market and solidifying Netflix's position as a major player in the region.





#### Netflix Expands Spanish-Language Content After Record Global Growth
Netflix Doubles Down on Spanish-Language Slate as Global Hits Drive Record International Growth

LOS ANGELES, CA – Netflix today announced a major expansion of its Spanish-language catalog, capitalizing on the unprecedented global demand for hits from Spain and Latin America. The streamer will launch 20 new Spanish-language originals in 2026 through production partnerships with studios in Mexico City, Madrid, and Bogotá. This strategic investment is fueled by the proven performance of its existing library; flagship titles like La Casa de Papel (Money Heist), Élite, and the recent spinoff Berlín have collectively amassed over 500 million hours of viewership worldwide, consistently ranking them among the most-watched non-English series on the platform. The success of these shows is not confined to Spanish-speaking countries, with data showing they regularly chart in the Top 10 in over 70 countries, including the United States, Germany, and India. This expansion underscores Netflix's core strategy of using locally-made stories with high production value to attract and retain a global audience, solidifying its position as the leader in international entertainment.


 


### **Acquire Data**

```{r}
if(!dir.exists(file.path("data", "mp01"))){
    dir.create(file.path("data", "mp01"), showWarnings=FALSE, recursive=TRUE)
}

GLOBAL_TOP_10_FILENAME <- file.path("data", "mp01", "global_top10_alltime.csv")

if(!file.exists(GLOBAL_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-global.tsv", 
                  destfile=GLOBAL_TOP_10_FILENAME)
}

COUNTRY_TOP_10_FILENAME <- file.path("data", "mp01", "country_top10_alltime.csv")

if(!file.exists(COUNTRY_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-countries.tsv", 
                  destfile=COUNTRY_TOP_10_FILENAME)
}
```

### **Data Import**
```{r}
if(!require("tidyverse")) install.packages("tidyverse")
library(readr)
library(dplyr)

GLOBAL_TOP_10 <- read_tsv(GLOBAL_TOP_10_FILENAME)

str(GLOBAL_TOP_10)
glimpse(GLOBAL_TOP_10)
```

### **Data Cleaning**
```{r}
GLOBAL_TOP_10 <- GLOBAL_TOP_10 |> mutate(season_title = if_else(season_title=="N/A",NA,season_title))
str(GLOBAL_TOP_10)
glimpse(GLOBAL_TOP_10)

```
### **Data Import**
```{r}
COUNTRY_TOP_10 <- read_tsv(COUNTRY_TOP_10_FILENAME, na = c("N/A", "NA"))
str(COUNTRY_TOP_10)
glimpse(COUNTRY_TOP_10)

# if (!require("DT")) install.packages('DT')
xfun::session_info('DT')
```

### **Initial Data Exploration**
```{r}
library(DT)
GLOBAL_TOP_10 |> 
    head(n=20) |>
    datatable(options=list(searching=FALSE, info=FALSE))

library(stringr)
format_titles <- function(df){
    colnames(df) <- str_replace_all(colnames(df), "_", " ") |> str_to_title()
    df
}

GLOBAL_TOP_10 |> 
    format_titles() |>
    head(n=20) |>
    datatable(options=list(searching=FALSE, info=FALSE)) |>
    formatRound(c('Weekly Hours Viewed', 'Weekly Views'))

GLOBAL_TOP_10 |> 
    mutate(`runtime_(minutes)` = round(60 * runtime)) |>
    select(-season_title, 
           -runtime) |>
    format_titles() |>
    head(n=20) |>
    datatable(options=list(searching=FALSE, info=FALSE)) |>
    formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
### **Exploratory Questions**
How many different countries does Netflix operate in?
```{r}
## 1

netflix_countries <- COUNTRY_TOP_10 |>
  distinct(country_name) |>
  n_distinct()

cat("Netflix operates in", netflix_countries, "countries.\n")
```
Which non-English-language film has spent the most cumulative weeks in the global top 10? How many weeks did it spend?
```{r}
## 2 
 GLOBAL_TOP_10 |>
    filter(category == 'Films (Non-English)') |>          
    filter(cumulative_weeks_in_top_10 == max(cumulative_weeks_in_top_10)) |>
    slice(which.max(cumulative_weeks_in_top_10)) |>
    select(show_title, cumulative_weeks_in_top_10)
```
What is the longest film (English or non-English) to have ever appeared in the Netflix global Top 10? How long is it in minutes?
```{r}
## 3
GLOBAL_TOP_10 |> 
    mutate(runtime_in_minutes = round(60 * runtime)) |>
    filter(category == "Films (English)" | category == "Films (Non-English)") |>
    select(category, show_title, runtime_in_minutes) |>
    slice_max(runtime_in_minutes)
```
For each of the four categories, what program has the most total hours of global viewership?
```{r}
##4

top_program <- GLOBAL_TOP_10 |>
  group_by(category, show_title) |>
  summarize(total_hours_viewed = sum(weekly_hours_viewed, na.rm = TRUE), .groups = 'drop') |>
  arrange(category, desc(total_hours_viewed)) |>
  group_by(category) |>
  slice(1) |>
  arrange(desc(total_hours_viewed)) |>
  slice(1)  

cat("The program with the most total hours of global viewership is", top_program$show_title, "in the", top_program$category, "category with", round(top_program$total_hours_viewed), "hours.\n")
```
Which TV show had the longest run in a country’s Top 10? How long was this run and in what country did it occur?
```{r}
##5
longest_country_run <- COUNTRY_TOP_10 |>
  filter(str_detect(category, "TV")) |>
  slice_max(cumulative_weeks_in_top_10, n = 1)

cat("The TV show with the longest run in a country's Top 10 is", longest_country_run$show_title, 
    "with", longest_country_run$cumulative_weeks_in_top_10, "weeks in", longest_country_run$country_name, ".\n")
```
Netflix provides over 200 weeks of service history for all but one country in our data set. Which country is this and when did Netflix cease operations in that country?
```{r}
country_weeks <- COUNTRY_TOP_10 |>
  group_by(country_name) |>
  summarise(
    total_weeks = n_distinct(week),
    first_week = min(week),
    last_week = max(week)
  ) |>
  arrange(total_weeks)


lowest_weeks_country <- country_weeks |>
  slice_min(total_weeks, n = 1)

print(lowest_weeks_country)
```
What is the total viewership of the TV show Squid Game? Note that there are three seasons total and we are looking for the total number of hours watched across all seasons.
```{r}
squid_game_total <- GLOBAL_TOP_10 |>
  filter(str_detect(tolower(show_title), "squid game")) |>
  summarise(total_hours_viewed = sum(weekly_hours_viewed, na.rm = TRUE))

cat("Squid Game total viewership:", squid_game_total$total_hours_viewed, "hours\n")
```


The movie Red Notice has a runtime of 1 hour and 58 minutes. Approximately how many views did it receive in 2021?
```{r}
red_notice_2021 <- GLOBAL_TOP_10 |>
  filter(str_detect(tolower(show_title), "red notice"),
         year(week) == 2021) |>
  summarise(
    total_hours = sum(weekly_hours_viewed, na.rm = TRUE),
    estimated_views = total_hours / (118/60) # 1小时58分钟 = 118分钟
  )

cat("Red Notice estimated views in 2021:", round(red_notice_2021$estimated_views), "views\n")
```

How many Films reached Number 1 in the US but did not originally debut there? That is, find films that first appeared on the Top 10 chart at, e.g., Number 4 but then became more popular and eventually hit Number 1? What is the most recent film to pull this off?
```{r}

us_films <- COUNTRY_TOP_10 |>
  filter(country_name == "United States",
         category == "Films") |>
  group_by(show_title) |>
  summarise(
    first_rank = min(weekly_rank, na.rm = TRUE),
    reached_number_one = any(weekly_rank == 1, na.rm = TRUE),
    last_week_in_chart = max(week, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(reached_number_one == TRUE, first_rank > 1)

total_films <- nrow(us_films)
most_recent_film <- us_films |>
  slice_max(last_week_in_chart, n = 1)

cat("Number of films that reached #1 in US but didn't debut there:", total_films, "\n")
cat("Most recent film to accomplish this:", most_recent_film$show_title, "\n")
```

Which TV show/season hit the top 10 in the most countries in its debut week? In how many countries did it chart?
```{r}
debut_week_performance <- COUNTRY_TOP_10 |>
  group_by(show_title, category) |>
  mutate(first_week_global = min(week, na.rm = TRUE)) |>
  filter(week == first_week_global) |>
  group_by(show_title, category, week) |>
  summarise(
    countries_charted = n_distinct(country_name),
    .groups = "drop"
  ) |>
  arrange(desc(countries_charted)) |>
  head(1)

print(debut_week_performance)
```


```{r}
#| include: false
1+1
```
--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`